---
title: "Regression Models Project"
author: "Harshad B."
date: "07/09/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Tasks

Take the *mtcars* data set and write up an analysis to answer their question using regression models and exploratory data analyses.

## Executive Summary

The data was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973–74 models).

A data frame with 32 observations on 11 (numeric) variables.

[, 1]	mpg	Miles/(US) gallon
[, 2]	cyl	Number of cylinders
[, 3]	disp	Displacement (cu.in.)
[, 4]	hp	Gross horsepower
[, 5]	drat	Rear axle ratio
[, 6]	wt	Weight (1000 lbs)
[, 7]	qsec	1/4 mile time
[, 8]	vs	Engine (0 = V-shaped, 1 = straight)
[, 9]	am	Transmission (0 = automatic, 1 = manual)
[,10]	gear	Number of forward gears
[,11]	carb	Number of carburetors

Loading the dataset:

```{r cars}
data(mtcars)
summary(mtcars)
```

```{r pressure}
mtcars$drat <- factor(mtcars$drat)
mtcars$cyl <- factor(mtcars$cyl)
mtcars$vs <- factor(mtcars$vs)
mtcars$gear <- factor(mtcars$gear)
mtcars$carb <- factor(mtcars$carb)
mtcars$am <- factor(mtcars$am,labels=c('Automatic','Manual'))
```

## Regression Model

``` {r}
full.model <- lm(mpg ~ ., data = mtcars)
best.model <- step(full.model, direction = "backward")
```

```{r}
summary(best.model)
```

A test to determine the difference in mpg between the automatic and manual transmissions.

``` {r}
t.test(mpg ~ am, data = mtcars)
```

Boxplot for the same

``` {r}
boxplot(mpg ~ am, data = mtcars, col = "blue", ylab = "miles per gallon")
```

The boxplot confirms the result.

## Conclusion

Therefore, in conclusion, it can be state that number of cylinders and displacement are good predictors of fuel efficiency, but transmission type is not.