---
title: "Statistical Interference Course Project"
author: "Harshad B."
date: "06/09/2020"
output: pdf_document
---

## Tasks to accomplish

1. Load the ToothGrowth data and perform some basic exploratory data analyses
2. Provide a basic summary of the data.
3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)
4. State your conclusions and the assumptions needed for your conclusions.

## Loading Libraries

```{r DataLoading}
library(ggplot2)
```

## Initialise

```{r Stuff}
# Load ToothGrowth data
data("ToothGrowth")

# Display a summary of the data
summary(ToothGrowth)

head(ToothGrowth)
```
```{r}
unique(ToothGrowth$len)
```
```{r}
unique(ToothGrowth$supp)
```
```{r}
unique(ToothGrowth$dose)
```
```{r}
# Convert dose to a factor
ToothGrowth$dose<-as.factor(ToothGrowth$dose)

# Plot tooth length ('len') vs. the dose amount ('dose'), broken out by supplement delivery method ('supp')
ggplot(aes(x=dose, y=len), data=ToothGrowth) + geom_boxplot(aes(fill=dose)) + xlab("Dose Amount") + ylab("Tooth Length") + facet_grid(~ supp) + ggtitle("Tooth Length vs. Dose Amount \nby Delivery Method") + 
     theme(plot.title = element_text(lineheight=.8, face="bold"))
```
```{r}
# Plot tooth length ('len') vs. supplement delivery method ('supp') broken out by the dose amount ('dose')
ggplot(aes(x=supp, y=len), data=ToothGrowth) + geom_boxplot(aes(fill=supp)) + xlab("Supplement Delivery") + ylab("Tooth Length") + facet_grid(~ dose) + ggtitle("Tooth Length vs. Delivery Method \nby Dose Amount") + 
     theme(plot.title = element_text(lineheight=.8, face="bold"))
```
```{r}
# run t-test
t.test(len~supp,data=ToothGrowth)
```
```{r}
# run t-test using dose amounts 0.5 and 1.0
ToothGrowth_sub <- subset(ToothGrowth, ToothGrowth$dose %in% c(1.0,0.5))
t.test(len~dose,data=ToothGrowth_sub)
```
```{r}
# run t-test using dose amounts 0.5 and 2.0
ToothGrowth_sub <- subset(ToothGrowth, ToothGrowth$dose %in% c(0.5,2.0))
t.test(len~dose,data=ToothGrowth_sub)
```
```{r}
# run t-test using dose amounts 1.0 and 2.0
ToothGrowth_sub <- subset(ToothGrowth, ToothGrowth$dose %in% c(1.0,2.0))
t.test(len~dose,data=ToothGrowth_sub)
```

As can be seen, the p-value of each test was essentially zero and the confidence interval of each test does not cross over zero (0).

Based on this result we can assume that the average tooth length increases with an inceasing dose, and therefore the null hypothesis can be rejected.

## Conclusions
Given the following assumptions:

1. The sample is representative of the population
2. The distribution of the sample means follows the Central Limit Theorem

In reviewing our t-test analysis from above, we can conclude that supplement delivery method has no effect on tooth growth/length, however increased dosages do result in increased tooth length.